---
title: "wk4-hw`kenwan`bayes"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rjags)
dataPath = "D:/datascience/bayesian"

suppressWarnings(source(paste(dataPath,"DBDA2Eprograms/DBDA2E-utilities.R",sep="/")))
library('rstan')

# test

fx <- inline::cxxfunction( signature(x = "integer", y = "numeric" ) , '
	return ScalarReal( INTEGER(x)[0] * REAL(y)[0] ) ;
' )
fx( 2L, 5 ) # should be 10

```

# Biomial output with two groups

Data are panel data containing columns: sex, race, age, education, state and y.
The output y is the binary version of the original data.
It reflects opinion (positive-1, negative-0) about President Barak Obama during his 2012 campaign for reelection.

Consider problem of identifying responses of two gender groups (column sex).

Decide what is the right model and analyze it using JAGS.I'll use stan instead.

Consider different prior distributions.

### Load in data

```{r}

myData = read.csv(paste0(dataPath,"/wk4/MScA_32014_BayesianMethods_CourseProjectData.csv"))
head(myData)   

dim(myData)

#positive
unique(myData$y)
sum(myData$y)

(dataList<-list(y=myData$y,Ntotal=length(myData$y)))
y<-myData$y
(Ntotal<-length(y))
```

# Simple binomial breaks down.

Let's examine then the impact of gender.

```{r}

unique(myData$sex)

#string

y = as.numeric(myData$y)
s = as.numeric(myData$sex) # ensures consecutive integer levels
# Do some checking that data make sense:

if ( any( y!=0 & y!=1 ) ) { stop("All y values must be 0 or 1.") }
Ntotal = length(y)
Nsubj = length(unique(s))

# Specify the data in a list, for later shipment to JAGS:
dataList = list(
  y = y ,
  s = s ,
  Ntotal = Ntotal ,
  Nsubj = Nsubj
  )

# THE MODEL.
modelString = "
data {
  int<lower=1> Nsubj ;
  int<lower=1> Ntotal ;
  int<lower=0,upper=1> y[Ntotal] ;
  int<lower=1> s[Ntotal] ; // notice Ntotal not Nsubj
}
parameters {
  real<lower=0,upper=1> theta[Nsubj] ; // individual prob correct
  real<lower=0,upper=1> omega ;        // group mode
  real<lower=0> kappaMinusTwo ;        // group concentration minus two
}
transformed parameters {
real<lower=0> kappa ;  
  kappa = kappaMinusTwo + 2 ;
}
model {
  omega ~ beta( 1 , 1 ) ;
  kappaMinusTwo ~ gamma( 0.01 , 0.01 ) ; // mean=1 , sd=10 (generic vague)
  // kappaMinusTwo ~ gamma( 1.105125 , 0.1051249 ) ;  # mode=1 , sd=10 
  theta ~ beta( omega*(kappa-2)+1 , (1-omega)*(kappa-2)+1 ) ; // vectorized
  for ( i in 1:Ntotal ) {
    y[i] ~ bernoulli( theta[s[i]] ) ;
  }
}
" # close quote for modelString



```

Initialize.

```{r}

# INTIALIZE THE CHAINS.
# Initial values of MCMC chains based on data:
initsList = function() {
  thetaInit = rep(0,Nsubj)
  for ( sIdx in 1:Nsubj ) { # for each subject
    includeRows = ( s == sIdx ) # identify rows of this subject
    yThisSubj = y[includeRows]  # extract data of this subject
    resampledY = sample( yThisSubj , replace=TRUE ) # resample
    thetaInit[sIdx] = sum(resampledY)/length(resampledY) 
  }
  thetaInit = 0.001+0.998*thetaInit # keep away from 0,1
  meanThetaInit = mean( thetaInit )
  kappaInit = 100 # lazy, start high and let burn-in find better value
  return( list( theta=thetaInit , omega=meanThetaInit , 
                kappaMinusTwo=kappaInit-2 ) )
}

# RUN THE CHAINS
parameters = c( "theta","omega","kappa") # The parameters to be monitored
burnInSteps = 500            # Number of steps to burn-in the chains
nChains = 4                  # nChains should be 2 or more for diagnostics 
numSavedSteps=50000
thinSteps=1

# Translate to C++ and compile to DSO:
stanDso <- stan_model( model_code=modelString ) 

```

Now run.

```{r}

# Get MC sample of posterior:
startTime = proc.time()
stanFit <- sampling( object=stanDso , 
                       data = dataList , 
                       #pars = parameters , # optional
                       chains = nChains ,
                       iter = ( ceiling(numSavedSteps/nChains)*thinSteps
                                +burnInSteps ) , 
                       warmup = burnInSteps , 
                       thin = thinSteps ,
                       init = initsList ) # optional  


stopTime = proc.time()
duration = stopTime - startTime
show(duration)

```

Convergence.

```{r}

show(stanFit)

plot(stanFit)

```

We can see that the theta values arevery well split, with nearly 0 sd in the samples. The same can not be said for the omega and kappa values.


```{r}
rstan::traceplot(stanFit,pars=c("omega","kappa"), ncol=1, inc_warmup=F)

pairs(stanFit, pars=c("omega","kappa"))

```

We can see that the oomega anmd kappa values are not centered well for omega.

```{r}

stan_hist(stanFit)

```

DIfferent visualization, but theta value histograms look like a strong estimate with low sd.

Autocorrelation.


```{r}

stan_ac(stanFit, separate_chains = T)

stan_diag(stanFit,information = "sample",chain=0)


```

MCMC trajecvtories


```{r}

OmegaKappa<-cbind(Omega=rstan::extract(stanFit,pars=c("omega","kappa"))$'omega',
                  Kappa=rstan::extract(stanFit,pars=c("omega","kappa"))$'kappa')

head(OmegaKappa)

Thetas<-rstan::extract(stanFit,pars=names(stanFit))
Thetas<-matrix(unlist(Thetas), ncol = 6, byrow = F)
colnames(Thetas)<-names(stanFit)
Thetas<-Thetas[,-(3:6)]
head(Thetas)

```

HDIS

```{r}
library(HDInterval)

hdi(as.vector(Thetas))
apply(Thetas,2,function(z) hdi(z))

```

