---
output: html_document
editor_options: 
  chunk_output_type: console
---

init

```{r}
library (faraway)
library(reshape)
```

data load

```{r}
data(femsmoke)

femdata = femsmoke
```

### joint dist:

```{r}
v <- c("smoke.yes", "smoke.no")
u <- c("dead.yes", "dead.no")
w <- c("age.1", "age.2", "age.3", "age.4", "age.5", "age.6", "age.7")

mat.femdata <- cast(femdata, smoker+dead ~ age, sum, value = 'y')
mat.dead.y <- mat.femdata[c(1,3), c(1, 3:9)]
mat.dead.n <- mat.femdata[c(2,4), c(1, 3:9)]
```

array

```{r}
rownames(mat.dead.y)<-rownames(mat.dead.y)<-v
mat.dead.y <- mat.dead.y[,c(2:8)]
rownames(mat.dead.n)<-rownames(mat.dead.n)<-v
mat.dead.n <- mat.dead.n[,c(2:8)]
```

reshape


```{r}
femdata_array<-array(rep(NA,28),
                       dim=c(2,7,2),
                       dimnames=list(paste("smoke",c("yes", "no"),
                                           sep="."),
                                     paste("age",1:7,sep="."),
                                     paste("dead",c("yes", "no"),sep=".")))

femdata_array[,,1] <- as.vector(unlist(mat.dead.y))
femdata_array[,,2] <- as.vector(unlist(mat.dead.n))
femdata_array
```


```{r}
N<-sum(femdata_array)
femdata.joint.p<-femdata_array/N

head(femdata.joint.p)

# check on sum
sum(femdata.joint.p)
```

### marginal distribution

marginal distribution for u_dead

```{r}
(uMarginal.dead <- apply(femdata.joint.p, 3, sum))

# check 
sum(uMarginal.dead)
```

marginal distribution for v_smoke

```{r}
(vMarginal.smoke <- apply(femdata.joint.p, 1, sum))
#check
sum(vMarginal.smoke)
```

marginal distribution for w_age.

```{r}
(wMarginal.age <- apply(femdata.joint.p, 2, sum))

sum(wMarginal.age)
```

### Create conditional distribution 

p(w,v|u="alive")=p(smoke,age|alive).

```{r}
(cond.v.w.given.uAlive <- femdata.joint.p[,,"dead.no"]/uMarginal.dead["dead.no"])

sum(cond.v.w.given.uAlive)
```

p(v|u="alive")=p(smoke|alive)

```{r}
(cond.v.given.uAlive <- apply(femdata.joint.p[,,"dead.no"], 1, sum)/uMarginal.dead["dead.no"])

sum(cond.v.given.uAlive)
```

p(w|u="alive",v="smoker")=p(age|alive,smoke)

```{r}
(cond.w.given.uAlive.vSmoke <- femdata.joint.p[,,"dead.no"]["smoke.yes",]/sum(femdata.joint.p[,,"dead.no"]["smoke.yes",]))

sum(cond.w.given.uAlive.vSmoke)
```

Compare the vectors p(w|v2,u1)p(v2|u1)p(u1) and p(w,v,u)[,v2,u1]

```{r}
rbind(uMarginal.dead["dead.no"]*cond.v.given.uAlive["smoke.yes"]*cond.w.given.uAlive.vSmoke,
      femdata.joint.p["smoke.yes",,"dead.no"])
```

### Let the marginal distribution for age group be p(w) estimated marginal distribution from the sample:

Given simulated age group, simulate variable v using conditional distribution p(v|w),i.e. using probabilities p(smoke.yes|age), p(smoke.no|age).

given simulated variables for age and for smoke, simulate mortality variable using distribution p(dead|v,w), p(alive|v,w).

Using the described procedure simulate outcomes for 100 participants. Use seed set.seed(284) for comparison.

Simulate age group

```{r}
set.seed(284)
sim.age <- c(rep(NA,100))
sim.age <- sapply(sim.age, function(z) sample(w,1,replace=T, prob = wMarginal.age))
```

Simulate variable v using conditional distribution p(v|w): p(smoke.yes|age), p(smoke.no|age).

```{r}
set.seed(284)
sim.v <- sapply(sim.age, function(z) sample(c("smoke.yes", "smoke.no"),1, replace=T,
                                            prob = vMarginal.smoke) )
```

Given simulated variables for age and smoke, simulate mortality
```{r}
set.seed(284)
sim.w <- apply(cbind(sim.age, sim.v),1,
               function(z) sample(c("dead.yes", "dead.no"), 1, replace=T,
                                                           prob=femdata.joint.p[z[2], z[1], ]))
```


```{r}
simulatedData <- cbind(sim.age, sim.v, sim.w)
rownames(simulatedData) <- 1:100
head(simulatedData, 25)
```


