---
title: "wk1-workshop"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

u<-c("u1","u2")
v<-c("v1","v2","v3")
w<-c("w1","w2","w3")
matr.u0<-paste("u0",outer(v,w,paste,sep=","),sep=",")
dim(matr.u0)<-c(3,3)
matr.u1<-paste("u1",outer(v,w,paste,sep=","),sep=",")
dim(matr.u1)<-c(3,3)
matr.u0

matr.u1

dataPath = "D:/datascience/bayesian/wk1/"

data3way<-read.csv(file=paste(dataPath,"3WayData.csv",sep="/"))
data3way = read.csv(file.choose())

head(data3way)

```

Data loaded. Now array

```{r}
mat.u0<-table(subset(data3way,u==0)[,1],subset(data3way,u==0)[,2])
mat.u1<-table(subset(data3way,u==1)[,1],subset(data3way,u==1)[,2])

mat.u0
mat.u1

idx.v1<-data3way$v==1
idx.w1<-data3way$w==1
idx.u1<-data3way$u==1
sum(idx.v1*idx.w1*idx.u1) #element (1,1) of mat.u1

idx.v2<-data3way$v==2
sum(idx.v2*idx.w1*idx.u1) #element (1,2) of mat.u1

idx.w2<-data3way$w==2
sum(idx.v1*idx.w2*idx.u1) #element (2,1) of mat.u1

colnames(mat.u1)<-colnames(mat.u0)<-c("v1","v2","v3")
rownames(mat.u1)<-rownames(mat.u0)<-c("w1","w2","w3")

data3way.array<-array(rep(NA,18),dim=c(3,3,2),dimnames=list(paste("w",1:3,sep=""),
                                                            paste("v",1:3,sep=""),
                                                            paste("u",0:1,sep="")))
data3way.array[,,1]<-mat.u0
data3way.array[,,2]<-mat.u1

data3way.array

N<-sum(data3way.array)
(data3way.array.p<-data3way.array/N)

uMarginal = apply(data3way.array.p,FUN = sum,MARGIN = 3)
vMarginal = apply(data3way.array.p,FUN = sum,MARGIN = 2)
wMarginal = apply(data3way.array.p,FUN = sum,MARGIN = 1)

uMarginal
vMarginal
wMarginal

```

So we pulled the marginal distributions for each one. Across both us.

Now the conditional portion.

Create conditional distribution p(w,v|u=1) as matrix cond.v.w.given.u1. So to do this we want to get the prob of w and v (any) given u1.

```{r}

cond.v.w.given.u1 = data3way.array.p[,,"u1"]/uMarginal["u1"]
cond.v.w.given.u1

```

Create conditional distribution p(v|u=1) as vector cond.v.given.u1

```{r}

cond.v.given.u1 = apply(data3way.array.p[,,"u1"],2,sum)/uMarginal["u1"]
cond.v.given.u1
```

Create conditional distribution p(w|v=2,u=1) as vector cond.w.given.u1.v2.

This should be p(w,v=2,u=1) / p(v=2 | u=1)*p(u=1)

```{r}
pwvu = data3way.array.p[,"v2","u1"]
uMarginal["u1"]

cond.w.given.u1.v2 = pwvu / (cond.v.given.u1["v2"]*uMarginal["u1"])
cond.w.given.u1.v2


```

SImulation

Let the marginal distribution for random variable u be Bernoulli with p(u=0)=0.55, p(u=1)=0.45.
Let conditional distributions for random variables (v|u=0) and (v|u=1), taking values 1,2,3 be

```{r}

(pCond.v.given.u0<-c(.7,.2,.1))
(pCond.v.given.u1<-c(.1,.2,.7))

p.given.u0.v1<-c(.3,.3,.4)
p.given.u0.v2<-c(.5,.3,.2)
p.given.u0.v3<-c(.6,.2,.2)
p.given.u1.v1<-c(.2,.3,.5)
p.given.u1.v2<-c(.2,.2,.6)
p.given.u1.v3<-c(.1,.7,.2)

```

Data generation

```{r}
set.seed(11)
sim.u = rbinom(n = 500,size = 1,prob = .45)

# generate data based on u value. 
sim.v = sapply(sim.u,
       function(z) sample(1:3,1,replace=TRUE,prob=ifelse(rep(z==1,3),pCond.v.given.u1,pCond.v.given.u0)))

simmed_res = data.frame(sim.u,sim.v)
head(simmed_res)

# BASED ON V  GENERATE
sim.w = apply(simmed_res,1,
              function(z) sample(1:3,1,replace=TRUE,
                                 prob=switch(3*z[1]+z[2],
                                             p.given.u0.v1,
                                             p.given.u0.v2,
                                             p.given.u0.v3,
                                             p.given.u1.v1,
                                             p.given.u1.v2,
                                             p.given.u1.v3)))

simmed_res["sim.w"] = sim.w

head(simmed_res)
```

EV

Uncond E[V]

```{r}

vMarginal %*% c(1,2,3)


```

Conditional E[v|u].

First, find conditional mean values E[v|u=u0]=E[v|u0] and E[v|u=u1]=E[v|u1].

The random variable (v|u0) takes values 1,2,3 with corresponding probabilities p(v=1|u0),p(v=2|u0),p(v=3|u0), given by the vector

```{r}


# E[v|u=u0] 

(cond.v.given.u0<-apply(data3way.array.p[,,"u0"],2,sum)/uMarginal["u0"])

# now value

exp.v.given.u0 = c(1,2,3) %*% cond.v.given.u0

# E[v|u=u1]

(cond.v.given.u1<-apply(data3way.array.p[,,"u1"],2,sum)/uMarginal["u1"])

exp.v.given.u1 = c(1,2,3) %*% cond.v.given.u1

# now total.

uMarginal %*% c(exp.v.given.u0,exp.v.given.u1)

```





