---
title: "wk4-hw`kenwan`bayes"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rjags)
dataPath = "D:/datascience/bayesian/"

suppressWarnings(source(paste(dataPath,"DBDA2Eprograms/DBDA2E-utilities.R",sep="/")))

```

# Biomial output with two groups

Data are panel data containing columns: sex, race, age, education, state and y.
The output y is the binary version of the original data.
It reflects opinion (positive-1, negative-0) about President Barak Obama during his 2012 campaign for reelection.

Consider problem of identifying responses of two gender groups (column sex).

Decide what is the right model and analyze it using JAGS.

Consider different prior distributions.

### Load in data

```{r}

myData = read.csv(paste0(dataPath,"wk4/MScA_32014_BayesianMethods_CourseProjectData.csv"))
head(myData)   

dim(myData)

 #positive
sum(myData$y)

(dataList<-list(y=myData$y,Ntotal=length(myData$y)))
y<-myData$y
(Ntotal<-length(y))
```

In aggregate only around 30% have positive sentiument oif obama in 2012.

Let's use a binomial model.

```{r}

# model init
modelString=" 
model {
  for (i in 1:length(Ntotal)) {
    y[i]~dbern(theta)
  }
  theta~dbeta(1,1)
}
"

writeLines(modelString,con=paste0(dataPath,"wk4/Tempmodel.txt"))


```

Initialize MCMC.

```{r}


MLE=sum(myData$y)/length(myData$y)
(init1<-MLE)
(init2<-MLE*(1+.01))
(init3<-MLE*(1-.01))

# fx
initsList<-function(){
  thetaInit<-sample(c(init1,init2,init3),1,replace=T)
  return(list(theta=thetaInit))
}

initsList()


```

Now JAGS

```{r}

jagsModel<-jags.model(file="Tempmodel.txt",data=dataList,n.chains=3,n.adapt=500)

```

Run


```{r}

update(jagsModel,n.iter=600)

codaSamples<-coda.samples(jagsModel,variable.names=c("theta"),n.iter=3334)
list.samplers(jagsModel)

head(codaSamples)

summary(codaSamples)
```

So the mean value has a pretty high std error. We can see in the theta values as well how spread the theta estimates can be.

More diagnostics.


```{r}

coda::traceplot(codaSamples)

densplot(codaSamples)

```

We can see in the density plot that we ahve no gaussian shape to theta. 

```{r}

autocorr.plot(codaSamples,ask=F)
effectiveSize(codaSamples)

``` 

Don't see autocvorrelation, though, so don't believe we havbe conmvergence issues.


```{r}

gelman.diag(codaSamples)
gelman.plot(codaSamples)


```

Gelman shrinkage suggests convergence was fine as well.


```{r}

lapply(codaSamples,mean)
sum(y)/Ntotal

# posterior densities vs theortical
(l<-min(unlist(codaSamples))-.05)
(h<-max(unlist(codaSamples))+.05)

histBreaks<-seq(l,h,by=.05)
postHist<-lapply(codaSamples,hist,breaks=histBreaks)

```

# We can see that our simple binomial breaks down.

Most likely we have two different groups (republican vs democratic). They have the same prior, but diufferent thetas.

However we don't have that information. Let's see if one based solely upon race would work.

```{r}

unique(myData$race)
unique(as.numeric(myData$race))
unique(myData$education)

#string
(s<-as.numeric(myData$race))
(Nsubj=length(unique(s)))
(dataList<-list(y=y,s=s,Ntotal=Ntotal,Nsubj=Nsubj))

# model

modelString = "
model {
  for ( i in 1:Ntotal ) {
    y[i] ~ dbern( theta[s[i]] )
  }
  for ( sIdx in 1:Nsubj ) {      # Different thetas from same prior
    theta[sIdx] ~ dbeta( 1 , 1 ) # N.B.: 2,2 prior; change as appropriate.
  }
}
" # close quote for modelString
writeLines( modelString , con="TEMP2model.txt" )


```

Initialize MCMC.

```{r}


MLE=sum(myData$y)/length(myData$y)
(init1<-MLE)
(init2<-MLE*(1+.01))
(init3<-MLE*(1-.01))

# fx
initsList<-function(){
  thetaInit<-sample(c(init1,init2,init3),1,replace=T)
  return(list(theta=thetaInit))
}

initsList()


```

Now JAGS

```{r}

jagsModel<-jags.model(file="TEMP2model.txt",data=dataList,n.chains=3,n.adapt=500)

```

Run


```{r}

update(jagsModel,n.iter=600)

codaSamples<-coda.samples(jagsModel,variable.names=c("theta"),n.iter=3334)
list.samplers(jagsModel)

head(codaSamples)

summary(codaSamples)
```

There we go! Look at the mean and sd for each. Naive SE is much lower e4 vs e2 in the first.

White (group1) has a very different mean
theta as does black.

```{r}

coda::traceplot(codaSamples)

densplot(codaSamples)

```

Shape is better now too.

```{r}

autocorr.plot(codaSamples,ask=F)
effectiveSize(codaSamples)

``` 

Don't see autocvorrelation, though, so don't believe we havbe conmvergence issues.


```{r}

gelman.diag(codaSamples)
gelman.plot(codaSamples)


```

Gelman shrinkage suggests convergence was fine as well.


```{r}
matrix(unlist(lapply(codaSamples,function(z) apply(z,2,mean))),ncol=3)

plot(density(codaSamples[[1]][,1]),xlim=c(0,1),ylim=c(0,150))
lines(density(codaSamples[[1]][,2]))
lines(density(codaSamples[[1]][,3]))
lines(density(codaSamples[[1]][,4]))


lines(density(codaSamples[[2]][,1]),col="orange")
lines(density(codaSamples[[2]][,2]),col="orange")
lines(density(codaSamples[[2]][,3]),col="orange")
lines(density(codaSamples[[2]][,4]),col="orange")

lines(density(codaSamples[[3]][,1]),col="blue")
lines(density(codaSamples[[3]][,2]),col="blue")
lines(density(codaSamples[[3]][,3]),col="blue")
lines(density(codaSamples[[3]][,4]),col="blue")

```

HDI


```{r}
(HDIofChainsHierarchi<-lapply(codaSamples,function(z) cbind(Theta1=HDIofMCMC(codaSamples[[1]][,1]),                                                   Theta2=HDIofMCMC(codaSamples[[1]][,2]),Theta3=HDIofMCMC(codaSamples[[1]][,3]),Theta4=HDIofMCMC(codaSamples[[1]][,4]))))


```


### lastly we'll apply different priors


```{r}
# model

modelString = "
model {
  for ( i in 1:Ntotal ) {
    y[i] ~ dbern( theta[s[i]] )
  }
  for ( sIdx in 1:Nsubj ) {      # Different thetas from same prior
    theta[sIdx] ~ dbeta( 2 , 2 ) # N.B.: 2,2 prior; change as appropriate.
  }
}
" # close quote for modelString
writeLines( modelString , con="TEMP3model.txt" )


```

Initialize MCMC. different MLE starts as well.

```{r}


MLE=sum(myData$y)/length(myData$y)
(init1<-MLE)
(init2<-MLE*(1+.5))
(init3<-MLE*(1-.5))

# fx
initsList<-function(){
  thetaInit<-sample(c(init1,init2,init3),1,replace=T)
  return(list(theta=thetaInit))
}

initsList()


```

Now JAGS

```{r}

jagsModel<-jags.model(file="TEMP3model.txt",data=dataList,n.chains=3,n.adapt=500)

```

Run


```{r}

update(jagsModel,n.iter=600)

codaSamples2<-coda.samples(jagsModel,variable.names=c("theta"),n.iter=3334)
list.samplers(jagsModel)

head(codaSamples2)

```

Let's compare the two runs: run one with 1,1 prior of beta distribution and the second with 2,2 and wider MLE inits.


```{r}

#run1
summary(codaSamples)
#run2
summary(codaSamples2)

```

Both iterations have no problem converging.

```{r}

coda::traceplot(codaSamples2)

densplot(codaSamples2)

autocorr.plot(codaSamples2,ask=F)
effectiveSize(codaSamples2)

``` 


Gelman shrinkage

```{r}

gelman.diag(codaSamples2)
gelman.plot(codaSamples2)


```

Let's plot them together.

```{r}
matrix(unlist(lapply(codaSamples2,function(z) apply(z,2,mean))),ncol=3)
s
plot(density(codaSamples2[[1]][,1]),xlim=c(0,1),ylim=c(0,125),main = "Density comparison of thetas, blue is second")
lines(density(codaSamples2[[1]][,2]))
lines(density(codaSamples2[[1]][,3]))
lines(density(codaSamples2[[1]][,4]))

lines(density(codaSamples[[1]][,1]),col="blue",lty="dotted")
lines(density(codaSamples[[1]][,2]),col="blue",lty="dotted")
lines(density(codaSamples[[1]][,3]),col="blue",lty="dotted")
lines(density(codaSamples[[1]][,4]),col="blue",lty="dotted")

```

### Discussion

We can see that the two runs ended up with nearly identical theta distributions regardless of prior and MLE start values. There is such a clear split based upon race that the posteriors converge quickly and separate neatly aside from latinos and "others".

