---
title: "bayes-wk2-workshop"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("rjags")
```

According to CDC, probability of breast cancer among women of age 50-60 years old is 2.28%.
Screening using mammography has sensitivity (i.e. probability of detecting the disease correctly) 75%-94%. Specificity of the test (i.e. probability of correct "no disease" conclusion) is 83%-98%. Assume the most optimistic parameters of the test (94% and 98%, correspondingly for sensitivity and specificity) What is the probability that randomly selected woman with positive test result has the disease?

p(tau1 | theta 1) = p(theta 1 | tau 1) * p(theta1) / (p(theta1 | tau1 ) * p(theta1) + p(theta1 | tau0) * p(theta0))

```{r}

#prior
p_theta1 = .0228
p_theta0 = 1-p_theta1

# sensitivity p(tau1 | theta1)
p_tau1_theta1 = .94
p_tau0_theta1 = 1 - p_tau1_theta1

# specificity p(tau0 | theta0)
p_tau0_theta0 = .98
p_tau1_theta0 = 1 - p_tau0_theta0

(p_theta1_tau1 = p_tau1_theta1 * p_theta1 / ((p_tau1_theta1 * p_theta1) + (p_tau1_theta0 * p_theta0)))

```

Assume that randomly selected woman tested positive and then retested with negative test result.
After both tests what is the probability of disease?

Prior is now p_theta1_tau1

Looking for p(theta1 | tau 0) 
  = (p(tau0 | theta 1)) * p(theta1) / (p(tau0 | theta1) * P(theta1) + p(tau0 | theta0) * p(theta0))
  
  
```{r}

pDisease = p_theta1_tau1
pnoDisease = 1-pDisease

(p_theta1_tau0 = (p_tau0_theta1 * pDisease / (p_tau0_theta1 * p_theta1 + p_tau0_theta0 * pnoDisease)))

```

#3.2

```{r}

suppressWarnings(library(faraway))
data(babyfood)
babyfood

(data<-babyfood[c(1,3),c(1:2,4)])

joint.dist<-data
joint.dist[1:2,1:2]<-joint.dist[1:2,1:2]/sum(joint.dist[1:2,1:2])
joint.dist


```

Marginal

```{r}

p_breast = sum(joint.dist[2,1:2])
p_disease = sum(joint.dist[,1])

p_breast_disease = joint.dist[2,1]/p_disease

```

p_disease_breast 

```{r}

p_disease_breast = p_breast_disease*p_disease / p_breast

```

#4

```{r}

#graphics.off()
getwd()
setwd("D:/datascience/bayesian/")

suppressWarnings(source("DBDA2Eprograms/BernGrid.R"))


```


```{r}

(Theta = seq( 0 , 1 , length=5 ))  # Sparse teeth for Theta.


pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
(pTheta = pTheta/sum(pTheta))      # Make pTheta sum to 1.0

Data = c(rep(0,0),rep(1,1))      # Single flip with 1 head

```


```{r}

# likelihood:

likelihood = Theta^Data*(1-Theta)^(1-Data)
post = pTheta*likelihood  

post = post/sum(post)
(posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="None" , showHDI=FALSE , showpD=FALSE ))
```

#4.5

```{r}

Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,3),rep(1,1))      # 25% heads, N=4

(likelihood=Theta^Data*(1-Theta)^4-sum(Data))
post = likelihood*pTheta

post=post/sum(post)

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=FALSE , showpD=FALSE )

```

#5 simulation of posterior

```{r}

likeli<-function(par,data){
  sdata<-sum(data)
  ldata<-length(data)
  return(par^sdata*(1-par)^(ldata-sdata))
}

Theta = seq( .00001 , 1 -.00001, length=1001 ) # Fine teeth for Theta.
pTheta = rep(1,length(Theta))      # Uniform (horizontal) shape for pTheta.
pTheta = pTheta/sum(pTheta)        # Make pTheta sum to 1.0



```

Now data generation and sample of theta from prior.

```{r}

set.seed(5)
(Data<-rbinom(5,size=1,prob=.84))

head(pTheta)

set.seed(15)
priorInd<-sample(1:length(Theta),500,replace = T)
priorSample<-cbind(Theta=Theta[priorInd],Prob=pTheta[priorInd])
priorSample<-rbind(priorSample,
                   c(head(Theta,1),head(pTheta,1)),
                   c(tail(Theta,1),tail(pTheta,1)))

likelihoodVector<-sapply(priorSample[,"Theta"],function(z) likeli(z,Data))
plot(priorSample[,"Theta"],likelihoodVector)

```

Now posterkior

```{r}

postVector<-priorSample[,"Prob"]*likelihoodVector
postVector<-postVector/sum(postVector)
plot(priorSample[,"Theta"],postVector)

postDistr<-approxfun(priorSample[,"Theta"],postVector,method="linear")
plot(priorSample[,"Theta"],postVector)
lines(Theta,postDistr(Theta),col="red",lwd=2)


```

Mode mean var
```{r}

postDistr1<-postDistr(Theta)
(mode1<-Theta[which.max(postDistr(Theta))])
(mean1<-Theta%*%postDistr(Theta)/sum(postDistr(Theta)))
(var1<-((Theta-mean1)^2)%*%postDistr(Theta)/sum(postDistr(Theta)))


```

replace 

```{r}

set.seed(25)
pTheta<-postDistr(Theta)/sum(postDistr(Theta))
(Data<-rbinom(5,size=1,prob=.84))

set.seed(35)
priorInd<-sample(1:length(Theta),500,replace = T)
priorSample<-cbind(Theta=Theta[priorInd],Prob=pTheta[priorInd])
priorSample<-rbind(priorSample,
                   c(head(Theta,1),head(pTheta,1)),
                   c(tail(Theta,1),tail(pTheta,1)))

likelihoodVector<-sapply(priorSample[,"Theta"],function(z) likeli(z,Data))
plot(priorSample[,"Theta"],likelihoodVector)


```a