---
title: "bayes-hw-wk2-kenwan"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(faraway)

```

## Wk2 - assignment. Posterior distribution for binomial.

Repeat calculations of Section 5 of the workshop:

Find posterior probability for binomial model with uniform prior and data. Use set.seed(81) for simulation of θ

Set likelihood function for binom.

```{r}

likeli<-function(par,data){
  sdata<-sum(data)
  ldata<-length(data)
  return(par^sdata*(1-par)^(ldata-sdata))
}

```

Generate theta. Uniform - 1000 samples.

```{r}

set.seed(81)
(data1<-rbinom(10,1,.71))

Theta = seq( .00001 , 1 -.00001, length=1001 ) # Fine teeth for Theta.
pTheta = rep(1,length(Theta))      # Uniform (horizontal) shape for pTheta.
pTheta = pTheta/sum(pTheta)        # Make pTheta sum to 1.0


```

Now data generation and sample of theta from prior.

```{r}

priorInd<-sample(1:length(Theta),500,replace = T)
priorSample<-cbind(Theta=Theta[priorInd],Prob=pTheta[priorInd])
priorSample<-rbind(priorSample,
                   c(head(Theta,1),head(pTheta,1)),
                   c(tail(Theta,1),tail(pTheta,1)))

likelihoodVector<-sapply(priorSample[,"Theta"],function(z) likeli(z,data1))
plot(priorSample[,"Theta"],likelihoodVector)

```

Now posterior.

```{r}

postVector<-priorSample[,"Prob"]*likelihoodVector
postVector<-postVector/sum(postVector)
plot(priorSample[,"Theta"],postVector)

postDistr<-approxfun(priorSample[,"Theta"],postVector,method="linear")
plot(priorSample[,"Theta"],postVector)
lines(Theta,postDistr(Theta),col="red",lwd=2)


```

### Now we have our posterior sample as well.

Mode mean var

```{r}

postDistr1<-postDistr(Theta)
(mode1<-Theta[which.max(postDistr(Theta))])
(mean1<-Theta%*%postDistr(Theta)/sum(postDistr(Theta)))
(var1<-((Theta-mean1)^2)%*%postDistr(Theta)/sum(postDistr(Theta)))


```

# Now data generation part 2.

```{r}

set.seed(97)
pTheta<-postDistr(Theta)/sum(postDistr(Theta))
# new data generation
(data2<-rbinom(10,1,.71))

priorInd<-sample(1:length(Theta),500,replace = T)
priorSample<-cbind(Theta=Theta[priorInd],Prob=pTheta[priorInd])
priorSample<-rbind(priorSample,
                   c(head(Theta,1),head(pTheta,1)),
                   c(tail(Theta,1),tail(pTheta,1)))

likelihoodVector<-sapply(priorSample[,"Theta"],function(z) likeli(z,data2))
plot(priorSample[,"Theta"],likelihoodVector)

```

New posterior.

```{r}

postVector<-priorSample[,"Prob"]*likelihoodVector
postVector<-postVector/sum(postVector)
plot(priorSample[,"Theta"],postVector)

postDistr<-approxfun(priorSample[,"Theta"],postVector,method="linear")
plot(priorSample[,"Theta"],postVector)
lines(Theta,postDistr(Theta),col="red",lwd=2)

postDistr2<-postDistr(Theta)
(mode2<-Theta[which.max(postDistr(Theta))])
(mean2<-Theta%*%postDistr(Theta)/sum(postDistr(Theta)))
(var2<-((Theta-mean2)^2)%*%postDistr(Theta)/sum(postDistr(Theta)))


```

# Now repeat for the reverse.

### Now data generation and sample of theta from prior using data2

```{r}

priorInd<-sample(1:length(Theta),500,replace = T)
priorSample<-cbind(Theta=Theta[priorInd],Prob=pTheta[priorInd])
priorSample<-rbind(priorSample,
                   c(head(Theta,1),head(pTheta,1)),
                   c(tail(Theta,1),tail(pTheta,1)))

likelihoodVector<-sapply(priorSample[,"Theta"],function(z) likeli(z,data2))
plot(priorSample[,"Theta"],likelihoodVector)

```

Now posterior.

```{r}

postVector<-priorSample[,"Prob"]*likelihoodVector
postVector<-postVector/sum(postVector)
plot(priorSample[,"Theta"],postVector)

postDistr<-approxfun(priorSample[,"Theta"],postVector,method="linear")
plot(priorSample[,"Theta"],postVector)
lines(Theta,postDistr(Theta),col="red",lwd=2)


```

Mode mean var

```{r}

postDistr3<-postDistr(Theta)
(mode3<-Theta[which.max(postDistr(Theta))])
(mean3<-Theta%*%postDistr(Theta)/sum(postDistr(Theta)))
(var3<-((Theta-mean3)^2)%*%postDistr(Theta)/sum(postDistr(Theta)))


```

# Reupdate with data1 now

```{r}

pTheta<-postDistr(Theta)/sum(postDistr(Theta))

priorInd<-sample(1:length(Theta),500,replace = T)
priorSample<-cbind(Theta=Theta[priorInd],Prob=pTheta[priorInd])
priorSample<-rbind(priorSample,
                   c(head(Theta,1),head(pTheta,1)),
                   c(tail(Theta,1),tail(pTheta,1)))

likelihoodVector<-sapply(priorSample[,"Theta"],function(z) likeli(z,data1))
plot(priorSample[,"Theta"],likelihoodVector)

```

New posterior.

```{r}

postVector<-priorSample[,"Prob"]*likelihoodVector
postVector<-postVector/sum(postVector)
plot(priorSample[,"Theta"],postVector)

postDistr<-approxfun(priorSample[,"Theta"],postVector,method="linear")
plot(priorSample[,"Theta"],postVector)
lines(Theta,postDistr(Theta),col="red",lwd=2)

postDistr4<-postDistr(Theta)
(mode4<-Theta[which.max(postDistr(Theta))])
(mean4<-Theta%*%postDistr(Theta)/sum(postDistr(Theta)))
(var4<-((Theta-mean4)^2)%*%postDistr(Theta)/sum(postDistr(Theta)))


```

# Comparison

```{r}

final_comp = rbind(data.frame(mode = mode2,mean = mean2,var = var2),data.frame(mode=mode4,mean = mean4, var = var4))

```